---
title: "Introduction to lt"
output:
  html:
    meta:
      css: ['@default', '@article', '@copy-button', '@heading-anchor', '@pages']
      js: ['@sidenotes', '@appendix', '@toc-highlight', '@copy-button', '@heading-anchor', '@pages']
    options:
      toc: true
      number_sections: true
vignette: >
  %\VignetteIndexEntry{Introduction to lt}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
---

The **lt** package provides a lightweight grammar of tables. Build a table by
declaring a JSON spec (titles, spanners, row groups, footnotes, formatters,
etc.); a tiny vanilla JavaScript runtime builds the HTML table from the spec on
page load. No sass, no V8, no htmlwidgets --- just base R and xfun.

A table in lt is built by composing independent operations:

1. **Data** -- the data frame to display.
2. **Header** -- title and subtitle.
3. **Spanners** -- labels above groups of columns.
4. **Row groups** -- partition rows by a column or manually.
5. **Footnotes** -- numbered references attached to any element.
6. **Notes** -- unnumbered footer text.
7. **Formatting** -- decimal places and thousands separators.
8. **Alignment** -- override default left/right alignment.

Each operation is added with the pipe operator `|>`, so building a table reads
naturally from left to right.

## A simple table

Every table starts with `lt()`, which accepts a data frame:

```{r}
library(lt)
lt(head(mtcars[, 1:6]))
```

## Title and subtitle

```{r}
lt(head(iris[, 1:4])) |>
  lt_header(title = "Iris Measurements", subtitle = "First six observations")
```

## Column spanners

Spanners group contiguous columns under a shared label:
```{r}
lt(head(iris)) |>
  lt_spanner("Sepal", c("Sepal.Length", "Sepal.Width")) |>
  lt_spanner("Petal", c("Petal.Length", "Petal.Width"))
```

## Row groups

Pass `row_group` to partition rows by a column's values:

```{r}
d = data.frame(
  Region = c("East", "East", "West", "West"),
  City = c("New York", "Boston", "Seattle", "Portland"),
  Pop = c(8336817, 675647, 737015, 652503)
)
lt(d, row_group = "Region") |>
  lt_format("Pop", big_mark = ",")
```

## Number formatting

```{r}
lt(head(mtcars[, 1:4])) |>
  lt_format(c("mpg", "disp"), decimals = 1) |>
  lt_format("drat", decimals = 2)
```

## Footnotes and notes

Footnotes are numbered automatically. Notes appear unnumbered in the footer:

```{r}
lt(head(mtcars[, 1:4])) |>
  lt_header(title = "Motor Trend Cars") |>
  lt_footnote("Source: 1974 Motor Trend US magazine.", "title") |>
  lt_footnote("Miles per US gallon.", "column", "mpg") |>
  lt_note("All measurements are for 1973--74 models.")
```

## Putting it all together

```{r}
d = data.frame(
  Group = c("Treatment", "Treatment", "Control", "Control"),
  Endpoint = c("Primary", "Secondary", "Primary", "Secondary"),
  Estimate = c(0.612, 0.789, 0.457, 0.543),
  CI_Lower = c(0.401, 0.568, 0.235, 0.321),
  CI_Upper = c(0.823, 1.010, 0.679, 0.765),
  P = c(0.001, 0.046, 0.123, 0.235)
)
lt(d, row_group = "Group", row_label = "Endpoint") |>
  lt_header(title = "Study Results", subtitle = "Efficacy endpoints") |>
  lt_spanner("95% CI", c("CI_Lower", "CI_Upper")) |>
  lt_format(c("Estimate", "CI_Lower", "CI_Upper"), decimals = 3) |>
  lt_format("P", decimals = 4) |>
  lt_footnote("Two-sided log-rank test.", "column", "P")
```
