---
title: "lt.js — JavaScript API Reference"
output:
  html:
    meta:
      css: ["@default", "@key-buttons"]
      js: ["@key-buttons"]
---

```{r assets, echo = FALSE, results = 'asis'}
# Pin jsdelivr URLs to the lt.js version in DESCRIPTION (not litedown's
# latest-tracking @-shortcuts). lt-plot loads before the core runtime.
cat(
  lt:::css_block(c('lt.css', 'lt-plot.css', 'lt-interactive.css'), inline = FALSE),
  lt:::js_block(c('lt-plot.js', 'lt.js', 'lt-interactive.js'), inline = FALSE),
  sep = '\n'
)
```

This document demonstrates the lt.js runtime directly from JavaScript.
Each example shows a JSON spec and renders the table with `LT.build()`.

## Including lt.js

Add the stylesheet and script to your HTML page:

``` html
<link rel="stylesheet" href="https://cdn.jsdelivr.net/npm/@xiee/utils/css/lt.min.css">
<script src="https://cdn.jsdelivr.net/npm/@xiee/utils/js/lt.min.js"></script>
```

Then call `LT.build(spec)` from an inline `<script>` wherever you want a
table to appear. The runtime renders the table immediately after the
calling script element. To build a table at any other time (e.g., when
the reader expands a row), call `LT.render(element, spec)` instead; see
[interactivity](#sec:interactive).

Interactive tables (sorting, search, filters, paging, resizing) need one
more stylesheet and script, loaded after the core ones:

``` html
<link rel="stylesheet" href="https://cdn.jsdelivr.net/npm/@xiee/utils/css/lt-interactive.min.css">
<script src="https://cdn.jsdelivr.net/npm/@xiee/utils/js/lt-interactive.min.js"></script>
```

Inline-graphics cells (error bars, sparklines, dot plots) live in a separate
graphics module. Load it **before** the core script, so its cell renderers
are registered by the time the runtime builds a table:

``` html
<link rel="stylesheet" href="https://cdn.jsdelivr.net/npm/@xiee/utils/css/lt-plot.min.css">
<script src="https://cdn.jsdelivr.net/npm/@xiee/utils/js/lt-plot.min.js"></script>
```

```{r, echo = FALSE}
iris5 = iris[c(1, 2, 51, 52, 101), ]  # span all three species
mtcars6 = head(mtcars[, 1:6], 6)
```

```{js, echo = FALSE}
// Shim: lt.js is deferred, so queue calls until it loads
if (!window.LT) window.LT = {
  q: [],
  build(spec) { this.q.push({s: document.currentScript, d: spec}); }
};
```

Define sample datasets as JavaScript objects (interpolated from R via
litedown's `fill` option):

```{js, fill = xfun::tojson}
const mtcars = `{ mtcars6 }`;
const iris = `{ iris5 }`;
```

## Basics

### Basic table

The simplest spec: just `data` (a column-oriented object).

```{js}
LT.build({ data: mtcars });
```

### Title and subtitle

```{js}
LT.build({
  data: iris,
  header: {
    title: "Iris Measurements",
    subtitle: "First five observations"
  }
});
```

### Column labels

By default, column names are cleaned for display by replacing `.` and
`_` with spaces. Set `auto_label: false` to show raw names, or use a
`label` op to override specific columns:

```{js}
// auto_label on by default: "Sepal.Length" displays as "Sepal Length"
LT.build({ data: iris });
```

```{js}
// Disable auto-labeling
LT.build({ data: iris, auto_label: false });
```

```{js}
// Explicit label overrides
LT.build({
  data: iris,
  ops: [
    { type: "label", labels: {
      "Sepal.Length": "Length (cm)", "Sepal.Width": "Width (cm)",
      "Petal.Length": "Length (cm)", "Petal.Width": "Width (cm)"
    }}
  ]
});
```

### Column alignment

By default, numeric columns are right-aligned. Override with an `align` op:

```{js}
LT.build({
  data: mtcars,
  ops: [
    { type: "align", columns: ["mpg", "cyl"], align: "center" }
  ]
});
```

## Formatting

### Auto-formatting

Numeric columns are automatically formatted. The following example shows
all behaviors at once: rounding to ~4 significant digits (adapting
decimal places to magnitude), thin-space thousands separator,
typographic minus (U+2212) for negatives, percentage detection (column
names matching `/%|[ _](pct|percent)$/i` are multiplied by 100 and
suffixed with "%"), and year-column skipping (columns matching
`/year/i` with 4-digit integers are left untouched).

```{js}
LT.build({
  data: {
    Year: [2020, 2021, 2022],
    Revenue: [1234567.89, 2345678.12, -3456.789],
    Margin_pct: [0.1234, 0.0567, 0.2345]
  }
});
```

#### Disabling auto-format

Set `auto_format: false` to show raw values. Compare with the table above:

```{js}
LT.build({
  data: {
    Year: [2020, 2021, 2022],
    Revenue: [1234567.89, 2345678.12, -3456.789],
    Margin_pct: [0.1234, 0.0567, 0.2345]
  },
  auto_format: false
});
```

### Number formatting (`fmt_number`)

Explicit control over decimals and thousands separators:

```{js}
LT.build({
  data: {
    Item: ["Widget", "Gadget", "Doohickey"],
    Revenue: [1234567.891, 987654.321, 246913.57],
    Margin: [0.1234, 0.0567, 0.2345]
  },
  ops: [
    { type: "fmt_number", columns: ["Revenue"], decimals: 0, big_mark: "," },
    { type: "fmt_number", columns: ["Margin"], decimals: 2, percent: true }
  ]
});
```

#### Significant digits

Use `sig_digits` instead of `decimals` to keep a fixed number of significant
digits, which suits columns spanning several orders of magnitude:

```{js}
LT.build({
  data: {
    Particle: ["Electron", "Proton", "Alpha"],
    Mass: [0.00054858, 1.00727647, 4.00150618]
  },
  ops: [
    { type: "fmt_number", columns: ["Mass"], sig_digits: 3 }
  ]
});
```

#### Percent without multiplication

Use `percent: "%"` when values are already in percent scale (no `*100`):

```{js}
LT.build({
  data: {
    Test: ["A", "B"],
    Rate: [78.5, 42.1]
  },
  ops: [
    { type: "fmt_number", columns: ["Rate"], decimals: 1, percent: "%" }
  ]
});
```

#### Prefix and suffix (currency, units)

Use `prefix` and `suffix` to prepend or append symbols:

```{js}
LT.build({
  data: {
    Item: ["Widget", "Gadget", "Doohickey"],
    Price: [1234.5, 678.9, 42.0],
    Weight: [1250, 340.5, 89.12]
  },
  ops: [
    { type: "fmt_number", columns: ["Price"], decimals: 2, big_mark: ",", prefix: "$" },
    { type: "fmt_number", columns: ["Weight"], decimals: 1, suffix: " kg" }
  ]
});
```

### Date formatting (`fmt_date`)

Format date values using JavaScript's native Date methods. Values can be
Date objects or ISO strings. The default calls `toLocaleDateString()`;
pass `method` to use a different Date method, or `locale`/`options` for
`Intl.DateTimeFormat` control.

**Note:** The formatted date may differ from the input depending on the
viewer's timezone. `new Date("2024-01-15")` is parsed as UTC midnight,
but `toLocaleDateString()` converts to local time — so `2024-01-15`
displays as `2024-01-14` at GMT-6. Pass `options: { timeZone: "UTC" }`
to avoid this.

```{js}
LT.build({
  data: {
    Event: ["Launch", "Update", "Release"],
    Date: [new Date("2024-01-15"), new Date("2024-06-30"), new Date("2024-12-01")]
  },
  ops: [
    { type: "fmt_date", columns: ["Date"] }
  ]
});
```

#### With locale and options

```{js}
LT.build({
  data: {
    Event: ["Launch", "Update"],
    Date: [new Date("2024-01-15"), new Date("2024-06-30")]
  },
  ops: [
    { type: "fmt_date", columns: ["Date"], locale: "zh-CN",
      options: { year: "numeric", month: "long", day: "numeric",
                 weekday: "long", timeZone: "UTC" } }
  ]
});
```

#### Using a specific Date method

```{js}
LT.build({
  data: {
    Event: ["Launch", "Update"],
    Date: [new Date("2024-01-15"), new Date("2024-06-30")]
  },
  ops: [
    { type: "fmt_date", columns: ["Date"], method: "toDateString" }
  ]
});
```

#### Datetime values

For datetime values, include time components in `options`:

```{js}
LT.build({
  data: {
    Event: ["Start", "End"],
    Time: [new Date("2024-01-15T09:30:00Z"), new Date("2024-01-15T17:45:00Z")]
  },
  ops: [
    { type: "fmt_date", columns: ["Time"], locale: "en-US",
      options: { year: "numeric", month: "short", day: "numeric",
                 hour: "2-digit", minute: "2-digit" } }
  ]
});
```

### Value substitution (`sub`)

Replace missing, zero, or small values with display text:

```{js}
LT.build({
  data: {
    Metric: ["HR", "p-value", "Events", "Rate"],
    Value: [0.62, 0.0003, 0, null]
  },
  ops: [
    { type: "fmt_number", columns: ["Value"], decimals: 2 },
    { type: "sub", columns: ["Value"], missing: "n/a", zero: "—", small: 0.001, small_text: "< 0.001" }
  ]
});
```

### Infinity values

`Inf` and `-Inf` (JSON: use `Infinity`/`-Infinity` — but since JSON doesn't
support `Infinity`, pass them as a JS expression):

```{js}
LT.build({
  data: {
    Metric: ["Lower bound", "Upper bound", "Rate", "Missing"],
    Value: [-Infinity, Infinity, 1.5, null]
  },
  ops: [
    { type: "fmt_number", columns: ["Value"], decimals: 2 }
  ]
});
```

## Column structure

### Column merging (`merge`)

Combine multiple columns into one display column. The `pattern` uses
`{1}`, `{2}`, etc. for column positions. Wrap sections in `<<` `>>` for
conditional NA handling.

```{js}
LT.build({
  data: {
    Endpoint: ["Primary", "Secondary", "Tertiary"],
    est: [0.61, 0.79, 0.45],
    ci_lo: [0.40, 0.57, null],
    ci_hi: [0.82, 1.01, null]
  },
  ops: [
    { type: "fmt_number", columns: ["est", "ci_lo", "ci_hi"], decimals: 2 },
    { type: "merge", columns: ["est", "ci_lo", "ci_hi"], pattern: "{1}<< ({2}, {3})>>" },
    { type: "label", labels: { est: "Estimate (95% CI)" } }
  ]
});
```

### Column spanners

Group contiguous columns under a shared header:

```{js}
LT.build({
  data: iris,
  spanners: [
    { label: "Sepal", columns: ["Sepal.Length", "Sepal.Width"] },
    { label: "Petal", columns: ["Petal.Length", "Petal.Width"] }
  ]
});
```

#### Auto-inferred spanners

Set `auto_span: true` to infer spanners from column names. Names are
split on the first `.` or `_`; contiguous columns sharing a prefix are
grouped, and labels are shortened to the suffix. Pass a custom regex
string (e.g., `"[.]"`) to change the separator.

```{js}
LT.build({ data: iris, auto_span: true });
```

### Column widths

```{js}
LT.build({
  data: mtcars,
  ops: [
    { type: "width", widths: { mpg: "10em", cyl: "5em", disp: "10em", hp: "8em" } }
  ]
});
```

### Table width

Set `table` on a `width` op to size the whole table. It can be combined
with column `widths`:

```{js}
LT.build({
  data: iris,
  ops: [
    { type: "width", table: "80%" }
  ]
});
```

```{js}
LT.build({
  data: iris,
  ops: [
    { type: "width", table: "100%", widths: { Species: "20em" } }
  ]
});
```

### Column reordering (`move`)

Move columns to the start or after a specific column:

```{js}
LT.build({
  data: iris,
  ops: [
    { type: "move", columns: ["Petal.Length", "Petal.Width"], after: null }
  ]
});
```

### Hiding columns (`hide`)

Drop columns from the table without dropping their data: the values still
travel in the spec, so they stay available to anything that reads the whole
row (e.g. an interactive table's row `detail`).

```{js}
LT.build({
  data: iris,
  ops: [
    { type: "hide", columns: ["Petal.Length", "Petal.Width"] }
  ]
});
```

## Inline plots

### Error bars (`errorbar`)

Render a value column's cells as a small inline SVG — a point at the
estimate with a bar (end caps included) from the lower to the upper bound —
on a scale shared across the column. `columns` lists the value column(s),
`lowers` and `uppers` the matching bound columns (one entry each per series);
the plot replaces the first value column's cells (hide the others). `min`/`max`
set the shared scale, `ref` draws a vertical reference line, and `width`/
`height` size each SVG. `axis: true` adds a shared axis in the footer (tick
marks + labels at "nice" round values, plus matching in-cell gridlines);
`axis_label` captions it. The SVG is built in the browser from the numbers, so
only the rows actually rendered (the current page, when interactive) are drawn.
This cell type comes from the graphics module (`lt-plot.js`), loaded before the
core script.

```{js}
LT.build({
  data: {
    Term: ["Age", "Dose", "Baseline", "Sex"],
    Est:  [0.12, -0.34, 0.45, -0.08],
    Lo:   [-0.05, -0.52, 0.21, -0.29],
    Hi:   [0.29, -0.16, 0.69, 0.13]
  },
  ops: [
    { type: "hide", columns: ["Lo", "Hi"] },
    { type: "errorbar", columns: ["Est"], lowers: ["Lo"], uppers: ["Hi"],
      min: -0.6, max: 0.7, ref: 0, axis: true, axis_label: "Effect (95% CI)" }
  ]
});
```

Several series stack in one cell: give `columns`/`lowers`/`uppers` one entry
per series, and `colors`/`labels` to key them in a footer legend. Each row then
draws one point-and-bar per series on the shared scale.

```{js}
LT.build({
  data: {
    Term: ["Headache", "Nausea", "Fatigue"],
    D1: [2.1, -1.4, 3.3], L1: [-1.0, -4.2, 0.1], U1: [5.2, 1.4, 6.5],
    D2: [-0.5, 0.8, 1.2], L2: [-3.6, -2.1, -1.9], U2: [2.6, 3.7, 4.3]
  },
  ops: [
    { type: "hide", columns: ["L1", "U1", "D2", "L2", "U2"] },
    { type: "errorbar",
      columns: ["D1", "D2"], lowers: ["L1", "L2"], uppers: ["U1", "U2"],
      colors: ["#5b8ff9", "#f6903d"], labels: ["Drug A", "Drug B"],
      min: -5, max: 7, ref: 0, axis: true, axis_label: "Risk difference (%)" }
  ]
});
```

### Sparklines (`sparkline`)

Draw a per-row series as a small inline line or bar chart. `columns` is either
a single list-column (each cell a numeric array) or several numeric columns
read across the row (the plot replaces the first column's cells; hide the
rest). `kind` is `"line"` (default) or `"bar"`; `min`/`max` set a shared
vertical scale (the default scales each row to its own range); `color` styles
the line or bars; `width`/`height` size each SVG. Like the error bar, it comes
from the graphics module (`lt-plot.js`) and is drawn in the browser.

```{js}
LT.build({
  data: {
    City: ["Oslo", "Cairo", "Lima"],
    Temp: [[-3, -2, 1, 6, 12, 16], [14, 15, 18, 23, 28, 31], [19, 19, 18, 16, 15, 15]],
    Rain: [[49, 36, 47, 41, 53, 60], [5, 4, 4, 1, 0, 0], [8, 4, 2, 1, 1, 1]]
  },
  ops: [
    { type: "sparkline", columns: ["Temp"] },
    { type: "sparkline", columns: ["Rain"], kind: "bar", color: "#4caf50" }
  ]
});
```

### Dot plots (`dotplot`)

Draw one dot per column on a scale shared across them, so a row with N columns
shows N dots laid out by value. `columns` lists the value columns (the plot
replaces the first column's cells; hide the rest). `colors` is an array of one
color per column — when present, a footer legend keys them, labeled by `labels`
(defaults to the column names). `min`/`max` set the shared scale; `axis: true`
adds the footer axis (and in-cell gridlines), `axis_label` captions it. Since
the cell shows every column, label its header with all the names joined (a
`label` op). Like the other plots, it comes from the graphics module
(`lt-plot.js`) and is drawn in the browser.

```{js}
LT.build({
  data: {
    Group:  ["Control", "Low", "High"],
    Before: [3.1, 3.4, 3.0],
    After:  [3.3, 4.1, 5.2]
  },
  ops: [
    { type: "hide", columns: ["After"] },
    { type: "label", labels: { Before: "Before / After" } },
    { type: "dotplot", columns: ["Before", "After"],
      colors: ["#999999", "#1a9641"], labels: ["Baseline", "Follow-up"],
      min: 2.5, max: 5.5, axis: true, axis_label: "Score" }
  ]
});
```

`stagger: true` spreads near-equal dots onto separate vertical tracks so
they do not overlap (the cell grows taller to fit one track per column):

```{js}
LT.build({
  data: {
    Group:  ["Control", "Low", "High"],
    Before: [3.1, 3.4, 3.0],
    After:  [3.2, 4.1, 5.2]
  },
  ops: [
    { type: "hide", columns: ["After"] },
    { type: "label", labels: { Before: "Before / After" } },
    { type: "dotplot", columns: ["Before", "After"], stagger: true,
      colors: ["#999999", "#1a9641"], labels: ["Baseline", "Follow-up"],
      min: 2.5, max: 5.5, axis: true, axis_label: "Score" }
  ]
});
```

## Row groups

### Group by column

Pass `row_group` as a single-element array to group by that column's
values (rendered as rowspan cells on the left by default):

```{js}
LT.build({
  data: {
    Region: ["East", "East", "West", "West", "West"],
    City: ["New York", "Boston", "Seattle", "Portland", "Denver"],
    Population: [8336817, 675647, 737015, 652503, 715522]
  },
  row_group: ["Region"]
});
```

### Auto separator rows for long labels

When `row_group` is a single-element array and the column is character
with any value longer than 20 characters, separator rows are used
automatically. Set `auto_sep: false` to force rowspan mode:

```{js}
LT.build({
  data: {
    Department: ["Cardiovascular Research", "Cardiovascular Research",
                 "Biostatistics", "Biostatistics"],
    Trial: ["ATLAS-2", "BEACON", "PRIME-1", "NEXUS"],
    N: [450, 320, 280, 510]
  },
  row_group: ["Department"]
});
```

```{js}
// Force rowspan mode by disabling auto-detection
LT.build({
  data: {
    Department: ["Cardiovascular Research", "Cardiovascular Research",
                 "Biostatistics", "Biostatistics"],
    Trial: ["ATLAS-2", "BEACON", "PRIME-1", "NEXUS"],
    N: [450, 320, 280, 510]
  },
  row_group: ["Department"],
  auto_sep: false
});
```

### Group by multiple columns (rowspan)

Pass multiple columns to render hierarchical rowspan cells:

```{js}
LT.build({
  data: {
    Region: ["East", "East", "East", "West", "West", "West"],
    State: ["NY", "NY", "MA", "WA", "WA", "OR"],
    City: ["New York", "Buffalo", "Boston", "Seattle", "Spokane", "Portland"],
    Population: [8336817, 278349, 675647, 737015, 228989, 652503]
  },
  row_group: ["Region", "State"]
});
```

### Separator-row style

Pass `row_group` as a string to always use separator rows:

```{js}
LT.build({
  data: {
    arm: ["Placebo", "Placebo", "Treatment", "Treatment"],
    stat: ["n", "Mean", "n", "Mean"],
    value: [30, 4.2, 31, 6.8]
  },
  row_group: "arm"
});
```

### Row group ordering

Reorder separator-row groups with a `group_order` op:

```{js}
LT.build({
  data: {
    arm: ["Placebo", "Placebo", "Treatment", "Treatment"],
    stat: ["n", "Mean", "n", "Mean"],
    value: ["30", "4.2", "31", "6.8"]
  },
  row_group: "arm",
  ops: [
    { type: "group_order", order: ["Treatment", "Placebo"] }
  ]
});
```

### Manual row groups

Define groups explicitly by row index:

```{js}
LT.build({
  data: mtcars,
  ops: [
    { type: "row_group", label: "First three", rows: [1, 2, 3] },
    { type: "row_group", label: "Last three", rows: [4, 5, 6] }
  ]
});
```

### Row indentation

Indent specific rows to show hierarchy:

```{js}
LT.build({
  data: {
    label: ["Any AE", "SOC: Cardiac", "Tachycardia", "Bradycardia", "SOC: GI", "Nausea"],
    n_pct: ["45 (67%)", "30 (45%)", "15 (22%)", "18 (27%)", "20 (30%)", "12 (18%)"]
  },
  ops: [
    { type: "indent", rows: [2, 5], level: 1 },
    { type: "indent", rows: [3, 4, 6], level: 2 }
  ]
});
```

## Styling

### Cell styling

Apply CSS to specific cells by column and/or row:

```{js}
LT.build({
  data: {
    Endpoint: ["Primary", "Secondary", "Exploratory"],
    HR: [0.62, 0.79, 0.91],
    P: [0.001, 0.042, 0.38]
  },
  ops: [
    { type: "fmt_number", columns: ["HR"], decimals: 2 },
    { type: "fmt_number", columns: ["P"], decimals: 3 },
    { type: "style", columns: ["P"], rows: [1, 2], css: "font-weight:bold;color:#06c" },
    { type: "style", columns: ["HR"], rows: [1], css: "background:#e8f4e8;border-bottom:2px solid #4a4" }
  ]
});
```

### Conditional styling

Use `test` with a JavaScript function to apply styles based on cell
values. The function receives the raw (pre-format) value and should
return `true` to apply. Use `class` to assign CSS classes instead of
inline `css`:

```{js}
LT.build({
  data: {
    Endpoint: ["Primary", "Secondary", "Exploratory"],
    HR: [0.62, 0.79, 1.05],
    P: [0.001, 0.042, 0.38]
  },
  ops: [
    { type: "fmt_number", columns: ["HR", "P"], decimals: 3 },
    { type: "style", columns: ["HR"], test: v => v < 1, class: "good" },
    { type: "style", columns: ["P"], test: v => v < 0.05, css: "font-weight:bold" }
  ]
});
```

```{css}
.lt-table .good { color: green; }
```

#### Highlighting missing values

Apply a class to all null cells across the table (omit `columns` to
target all):

```{js}
LT.build({
  data: {
    arm: ["Treatment", "Control", "Treatment"],
    n: [30, null, 28],
    response: [0.67, null, 0.71]
  },
  ops: [
    { type: "style", test: v => v == null, class: "na" }
  ]
});
```

```{css}
.lt-table .na { background: #fee; }
```

## HTML content

By default every cell value and text field is HTML-escaped, so characters
like `<`, `>`, and `&` render as literal text. Two mechanisms opt into raw
HTML:

- **Body cells:** set `html_cols` to `true` (all columns) or an array of
  column names. Cells in those columns are emitted verbatim.
- **Text fields** (title, subtitle, column labels, spanner labels, footnote
  and note text): pass a **one-element array** instead of a string. A plain
  string is escaped; `["<b>x</b>"]` is emitted as raw HTML. (This is what the
  R side's `I()` wrapper serializes to.)

```{js}
LT.build({
  data: {
    Package: [
      "<a href='https://cran.r-project.org/package=lt'>lt</a>",
      "<a href='https://cran.r-project.org/package=xfun'>xfun</a>"
    ],
    Status: ["<b style='color:#090'>stable</b>", "<i>active</i>"]
  },
  html_cols: ["Package", "Status"],
  header: { title: ["Packages <sup>&#9733;</sup>"] },
  ops: [
    { type: "label", labels: { Package: ["Name<sup>*</sup>"] } }
  ],
  notes: [["See <a href='https://yihui.org/'>the author</a>."]]
});
```

## Footnotes and notes

### Footnotes

Footnotes are de-duplicated by text and numbered automatically. Each
footnote's `location.type` selects a region: `title` (with
`group: "title"` or `group: "subtitle"`), `column_labels`,
`column_spanners`, `row_groups`, or `body`. The example below attaches a
footnote to every location:

```{js}
LT.build({
  data: {
    Arm: ["Treatment", "Treatment", "Control", "Control"],
    Stat: ["n", "Mean", "n", "Mean"],
    Value: [31, 6.8, 30, 4.2],
    SE: [null, 0.42, null, 0.38]
  },
  row_group: "Arm",
  header: { title: "Study Results", subtitle: "Two arms" },
  spanners: [{ label: "Summary", columns: ["Value", "SE"] }],
  footnotes: [
    // title and subtitle share type "title", distinguished by group
    { text: "Source: simulated trial data.", location: { type: "title", group: "title" } },
    { text: "Intention-to-treat population.", location: { type: "title", group: "subtitle" } },
    // a column label
    { text: "Standard error of the mean.", location: { type: "column_labels", columns: ["SE"] } },
    // a spanner label
    { text: "Per-arm summary statistics.", location: { type: "column_spanners", spanners: ["Summary"] } },
    // a row-group label
    { text: "Received active drug.", location: { type: "row_groups", match: "exact", values: ["Treatment"] } },
    // specific body cells (columns + 1-based rows)
    { text: "SE undefined for counts.", location: { type: "body", columns: ["SE"], rows: [1, 3] } }
  ]
});
```

### Source notes

Notes appear below footnotes in the footer:

```{js}
LT.build({
  data: mtcars,
  notes: ["Data from the 1974 Motor Trend US magazine."]
});
```

## Complete example

Combining multiple features in a single spec:

```{js}
LT.build({
  data: {
    Group: ["Treatment", "Treatment", "Control", "Control"],
    Endpoint: ["Primary", "Secondary", "Primary", "Secondary"],
    Estimate: [0.6123, 0.7891, 0.4567, 0.5432],
    CI_Lower: [0.4012, 0.5678, 0.2345, 0.321],
    CI_Upper: [0.8234, 1.0104, 0.6789, 0.7654],
    P_Value: [0.0012, 0.0456, 0.1234, 0.2345]
  },
  row_group: "Group",
  header: {
    title: "Study Results",
    subtitle: "Primary and secondary endpoints"
  },
  spanners: [
    { label: "95% CI", columns: ["CI_Lower", "CI_Upper"] }
  ],
  ops: [
    { type: "fmt_number", columns: ["Estimate", "CI_Lower", "CI_Upper"], decimals: 3 },
    { type: "fmt_number", columns: ["P_Value"], decimals: 4 },
    { type: "style", columns: ["P_Value"], rows: [1, 3], css: "font-weight:bold" }
  ],
  footnotes: [
    { text: "Two-sided p-value from log-rank test.", location: { type: "column_labels", columns: ["P_Value"] } }
  ]
});
```

## Interactivity {#sec:interactive}

### Raw values

When cell values are formatted, the original values are available in
tooltips (hover over a cell to see). You can also `Alt + Click` on a
table to reveal all raw values at once, or `Alt + Double-Click` to
toggle raw values for all tables on the page.

### The `interactive` field

With `lt-interactive.js` on the page, a spec that carries an
`interactive` object gets the requested controls. Every field is
optional:

- `sort`: click-to-sort headers, ascending then descending then
  unsorted (`false` to turn off); shift-click adds a column as a
  tie-breaker. An array requests an initial sort, each element a column
  name string, `"-name"` for descending (or a `{ col, dir }` object);
- `search`: a search box above the table, matching any cell (`false` to
  turn off);
- `filter`: per-column filters, as `{ default?, cols? }`. `default` is a
  spec applied to every visible column left unnamed; `cols` maps a column
  to its own spec. A spec is `true` for a plain box (matching that column
  the same way as the search), or `{ type: "select" }` / `{ type: "range" }`
  / `{ type: "checklist" }` for a value dropdown, a two-thumb slider, or a
  popover of checkboxes (choices / ends read from the column data, or set
  explicitly with `choices` / `min`+`max`; `selected` seeds the dropdown's
  value or the checklist's checked set). A `"select"` keeps the one chosen
  value, a `"checklist"` keeps every checked value (all checked = no
  filter). So `{ default: true }` boxes every column, `{ cols: { mpg: true
  } }` boxes one, and `{ cols: { cyl: { type: "select" } } }` gives `cyl` a
  dropdown;
- `pager`: the page sizes to offer, as an array whose first element is
  the initial size; `0` means every row on one page. Omit the field for
  the default sizes, or set it to `false` to show all rows with no
  pager;
- `resize`: drag a column's right edge to resize it (double-click that
  edge to fit the column to its content);
- `hide`: a column-visibility menu. An eye button at the start of the
  search row opens a checklist of every column, listed by its displayed
  label; unchecking one hides it (header and cells), re-checking restores
  it. `true` starts with all columns shown, `{ hidden: [...] }` starts
  the named columns hidden (all are still listed, so any can be toggled);
- `detail`: row detail (drill-down). An expand caret is added to each row,
  and a table is built below it on first expand. Give either an array of
  column names (the detail is a one-row table of those columns' displayed
  values, including ones hidden with a `hide` op) or a function `(row,
  index, display) => spec` (or the name of a global one) that returns the
  spec. `row` has the raw values and `display` the formatted text, each
  keyed by column name and covering every column. The returned spec is an
  ordinary table spec, so give it its own `interactive` field to make the
  detail table sortable/filterable/paged too;
- `download`: a button in the control bar that downloads the current view
  (every row the filters and search keep, in sort order, across all pages) as
  a CSV file, using the displayed cell text. `true` uses a default file name;
  a string sets the file name.

A term is matched case-insensitively against the displayed text, with a
leading `!` to negate it; a term that mentions `x` (e.g., `x > 20`) is
evaluated as a JavaScript expression against the raw cell values
instead.

```{js}
LT.build({
  data: mtcars,
  interactive: { filter: { cols: { mpg: true } }, pager: [3, 6, 0], resize: true }
});
```

Typed filters — a dropdown on `cyl`, a range slider on `wt`, a checklist on
`hp` — with a plain box as the default for the other columns. A typed filter
on a *hidden* column has no header to sit under, so it moves into the control
bar as a labelled chip (clicking anywhere on the chip opens its control):
here `gear` is hidden yet still filterable through its chip.

```{js}
LT.build({
  data: mtcars,
  interactive: {
    search: false, pager: 6,
    hide: { hidden: ["gear"] },
    filter: { default: true, cols: {
      cyl: { type: "select" }, wt: { type: "range" }, hp: { type: "checklist" },
      gear: { type: "select" }
    } }
  }
});
```

Open the eye menu to toggle columns; `hp` and `wt` start hidden here:

```{js}
LT.build({
  data: mtcars,
  interactive: { hide: { hidden: ["hp", "wt"] }, pager: 5 }
});
```

A download button writes the current view to a CSV file (filter or sort
first, and the file follows):

```{js}
LT.build({
  data: mtcars,
  interactive: { download: "cars.csv", filter: { default: true }, pager: 5 }
});
```

Row structure is honored: sorting and filtering happen *within* each
group or subtree, never across it. A `row_group` op sorts and filters
within each group — a separator group (a string column) whose rows all
filter out drops its header, while a rowspan group (an array of columns)
reorders its blocks when its header is clicked and stays nested with
several columns. An `indent` op sorts siblings under each parent and
keeps a match's ancestors visible. Spanners, styles, and footnotes are
no obstacle.

```{js}
LT.build({
  data: mtcars,
  row_group: ["cyl"],
  interactive: { pager: [8] }
});
```

### Rendering on demand

`LT.build()` has to run while the page is parsing, since it renders the
table where its `<script>` sits. `LT.render(element, spec)` has no such
constraint: it renders the spec into a container element whenever it is
called, returning the table element. Interactive tables built this way
are enhanced like any other, which is how a table can be built lazily —
for instance, only when the reader asks for it:

```{js}
const box = document.createElement("div");
const btn = document.createElement("button");
btn.textContent = "Build a table";
btn.onclick = () => {
  LT.render(box, { data: iris, interactive: { pager: [3] } });
  btn.disabled = true;
};
document.currentScript.after(box);
document.currentScript.after(btn);
```

### Driving a table from outside

Every enhanced table carries a small controller at `el._lt`, so widgets
outside the table — a dropdown, a range slider — can filter it without
using the built-in boxes:

- `el._lt.filter(id, func)` registers a predicate `func(row)` under `id`,
  where `row` is the row's raw values keyed by column; the rows kept are
  those for which `func` returns a truthy value. Call it again with the same
  `id` to replace that predicate, or with a `null` `func` to remove it. The
  predicates from every id — together with the built-in column filters and
  the search — combine with AND.
- `el._lt.refresh()` re-renders the body; `filter()` calls it for you, so
  you rarely need it directly.
- `el._lt.view()` returns the current view as 1-based original row indices
  (every row the filters and search keep, in sort order, across all pages) —
  what a "download what's shown" button reads.
- `el._lt.spec` and `el._lt.state` are exposed for reading. A widget reads
  `el._lt.spec.data[col]` for a column's raw values to build its choices
  (the unique categories for a dropdown, the min/max for a slider).

Predicates live in the table's state, so they survive sorting and paging.
Here an outside `<select>` filters the table by `Species`:

```{js}
// An outside widget wires up once the page (and lt.js) has loaded.
const host = document.createElement("div");
document.currentScript.after(host);
addEventListener("load", () => {
  const t = LT.render(host, { data: iris, interactive: { search: false, pager: [5] } });
  const sel = document.createElement("select");
  sel.innerHTML = '<option value="">All species</option>' +
    [...new Set(t._lt.spec.data.Species)].map(s => "<option>" + s + "</option>").join("");
  sel.onchange = () => t._lt.filter(
    "species", sel.value ? row => row.Species === sel.value : null
  );
  t.before(sel);
});
```
