---
title: "lt vs gt — gsDesign2 examples"
---

This document re-runs a few of the `as_gt()` examples from
[gsDesign2](https://github.com/Merck/gsDesign2)'s documentation, but builds
the tables with `lt` instead of `gt`. The first goal of `lt` is to be a
drop-in lightweight replacement for `gt` in gsDesign2 outputs.

Each chunk below is independent. Because `record_print.lt_tbl` always
emits the lt CSS+JS runtime (no knitr coupling), the runtime is included
once per chunk; the runtime self-guards (`if (root.LT) return`) so the
duplicate inclusions are no-ops at execution time.

```{r setup}
library(lt)
suppressMessages(library(gsDesign2))

enroll_rate = define_enroll_rate(duration = 18, rate = 20)
fail_rate = define_fail_rate(
  duration = c(4, 100), fail_rate = log(2) / 12,
  dropout_rate = .001, hr = c(1, .6)
)
study_duration = 36
alpha = 0.025
beta = 0.1
```

## Fixed design — AHR method

Equivalent of `fixed_design_ahr() |> summary() |> as_gt()`.

```{r fixed-ahr}
s = fixed_design_ahr(
  alpha = alpha, power = 1 - beta,
  enroll_rate = enroll_rate, fail_rate = fail_rate,
  study_duration = study_duration, ratio = 1
) |> summary()

s |>
  lt() |>
  lt_header(title = attr(s, "title")) |>
  lt_footnote(attr(s, "footnote"), "title") |>
  lt_format(c("N", "Events", "AHR", "Bound", "Power"), decimals = 4)
```

## Fixed design — Fleming-Harrington

Equivalent of `fixed_design_fh() |> summary() |> as_gt()`.

```{r fixed-fh}
s = fixed_design_fh(
  alpha = alpha, power = 1 - beta,
  enroll_rate = enroll_rate, fail_rate = fail_rate,
  study_duration = study_duration, ratio = 1
) |> summary()

s |>
  lt() |>
  lt_header(title = attr(s, "title")) |>
  lt_footnote(attr(s, "footnote"), "title") |>
  lt_format(c("N", "Events", "Bound", "Power"), decimals = 4)
```

## Group sequential — `gs_power_ahr`

Equivalent of `gs_power_ahr(lpar = ...) |> summary() |> as_gt()`. The long
"Analysis: N Time: ... AHR: ..." label is a natural row-group header
(gt does the same via `gt::gt(groupname_col = "Analysis", rowname_col =
"Bound")`).

```{r gs-power-ahr}
s = gs_power_ahr(lpar = list(sf = gsDesign::sfLDOF, total_spend = 0.1)) |>
  summary()

s |>
  lt(row_group = "Analysis", row_label = "Bound") |>
  lt_header(
    title = "Bound summary for AHR design",
    subtitle = "AHR approximations of ~HR at bound"
  ) |>
  lt_spanner(
    label = "Cumulative boundary crossing probability",
    columns = c("Alternate hypothesis", "Null hypothesis")
  ) |>
  lt_footnote(
    "Approximate hazard ratio to cross bound.",
    "column", "~HR at bound"
  ) |>
  lt_footnote(
    "One-sided p-value for experimental vs control treatment.",
    "column", "Nominal p"
  ) |>
  lt_format(
    c("Z", "~HR at bound", "Nominal p", "Alternate hypothesis", "Null hypothesis"),
    decimals = 4
  )
```

## Group sequential — `gs_design_ahr`

Equivalent of `gs_design_ahr() |> summary() |> as_gt()`.

```{r gs-design-ahr}
s = gs_design_ahr() |> summary()

s |>
  lt(row_group = "Analysis", row_label = "Bound") |>
  lt_header(
    title = "Bound summary for AHR design",
    subtitle = "AHR approximations of ~HR at bound"
  ) |>
  lt_spanner(
    label = "Cumulative boundary crossing probability",
    columns = c("Alternate hypothesis", "Null hypothesis")
  ) |>
  lt_format(
    c("Z", "~HR at bound", "Nominal p", "Alternate hypothesis", "Null hypothesis"),
    decimals = 4
  )
```

## What's missing vs `as_gt()`

A direct `as_lt()` wrapper analogous to `as_gt.gs_design_summary` (with
`title=`, `subtitle=`, `colname_spanner=`, `colname_spannersub=`,
`footnote=`, `display_bound=`, `display_columns=`, `display_inf_bound=`
arguments) is the obvious next step once the lt vocabulary settles. For
now the chunks above show that the gt → lt mapping is mechanical:

| gt                                | lt                                  |
|-----------------------------------|-------------------------------------|
| `gt::gt(groupname_col=, rowname_col=)` | `lt(row_group=, row_label=)`   |
| `gt::tab_header(title=, subtitle=)`    | `lt_header(title=, subtitle=)` |
| `gt::tab_spanner(label=, columns=)`    | `lt_spanner(label=, columns=)` |
| `gt::tab_footnote(footnote=, locations=cells_*())` | `lt_footnote(text, where, columns)` |
| `gt::tab_source_note(source_note=)`    | `lt_note(text)`                |
| `gt::fmt_number(columns=, decimals=)`  | `lt_format(columns, decimals)`    |
