---
title: Row Groups
---

```{r}
library(lt)
```

## Auto-detected row groups

Pass `row_group` to `lt()` to partition rows by a column's values. The
column is removed from the body and its values become group headers:

```{r}
d = data.frame(
  Region = c("East", "East", "West", "West", "West"),
  City = c("New York", "Boston", "Seattle", "Portland", "Denver"),
  Population = c(8336817, 675647, 737015, 652503, 715522)
)
lt(d, row_group = "Region") |>
  lt_header(title = "US Cities by Region") |>
  lt_format("Population", big_mark = ",")
```

## Row groups with a stub

Use `row_label` to designate a column as the row "stub" (left-aligned label
column with no header):

```{r}
d = data.frame(
  Category = c("Fruits", "Fruits", "Vegetables", "Vegetables"),
  Item = c("Apple", "Banana", "Carrot", "Broccoli"),
  Calories = c(95, 105, 25, 55),
  Fiber_g = c(4.4, 3.1, 2.8, 5.1)
)
lt(d, row_group = "Category", row_label = "Item") |>
  lt_header(title = "Nutritional Information") |>
  lt_format("Fiber_g", decimals = 1)
```

## Manual row groups

Use `lt_group()` when you need explicit control over which rows belong to
each group:

```{r}
lt(head(mtcars[, 1:4])) |>
  lt_group("First three", 1:3) |>
  lt_group("Last three", 4:6)
```
